Canonical rings of rational divisors on projective space and Hirzebruch surfaces

### 2015-07-16 surfaces-write-up/surface-write-up.tex
edits to projective space section

\ssec{Notation}
\begin{convention}
Write
\begin{align*}
	D = \sum_{i=1}^{n}\alpha_i D_i.
\end{align*}
where $\alpha_i \in \bq$. \todo{Peter: What are the $D_i$'s.}
\end{convention}

@@

\begin{proof}
By definition of $H^0(\bp^m,dD)$, functions satisfying conditions 
~\eqref{align:pm-span} lie in $H^0(\bp^m,dD)$. To complete the proof, it suffices to check functions satisfying conditions ~\eqref{align:pm-basis} form a basis of $H^0(\bp^m,dD)$. Note that there $\binom{m+ \lfloor dD \rfloor }{m}$ functions satisfying condition ~\eqref{align:pm-basis}. However we know $h^0(\bp^m,dD) = \binom{m+ \lfloor dD \rfloor }{m},$ so it suffices to show that those functions satisfying condition ~\eqref{align:pm-basis} are independent. This follows from the assumption that $f_0,\ldots, f_n$ form a basis of degree 1 rational functions, and so degree $\lfloor d \deg D \rfloor $ monomials in $f_0,\ldots, f_n$ form a basis of degree $\lfloor d \deg D \rfloor $ rational functions.
\end{proof}

We first restrict attention to the case $n=1$, so $D = \alpha D_i$.

\ssec{Effective Divisors of Hypersurfaces on Projective Space}
\label{ssec:proj-one-point}
\todo{intro}

@@

\begin{rem}
\label{rem:proj-one-point-ind}
The method of proof for Theorem ~\ref{thm:proj-one-point}
can be immediately generalized to work inductively. If $D = D' + H$
for some hyperplane $H$, then we can use the proof above to show
the canonical ring $R_D$ is generated over $R_D'$ by the same
generators and with the same relations given in Theorem
~\ref{thm:proj-one-point}.
\end{rem}

\begin{thm}
\label{thm:proj-effective}
Let $X = \bp^m$ with $\{x_0, \ldots, x_m\}$ a coordinate system for
$\bp^m$. Let $D = \sum_{i = 1}^{n} \alpha_i C_i$, with $\alpha_i =
\frac{p_i}{q_i} \in \bq$ and each $C_i$ a hypersurface of
$\bp^m$, be an effective divisor on $\bp^m$.

@@

\begin{proof}
We derive the result from the following inductive argument.
If $n = 0$, i.e.~ $D = 0$, then we are done.
Now, we look to inductively add hypersurfaces.
If $D' = \sum_{i = 0}^{n} \alpha_i C_i$ is some $\bq$-divisor
on $\bp^m$, then we look at $D = D' + C$ where $C$ is a
hypersurface of degree $\pdeg$. Now use the Veronese mapping
to view $D$ as a divisor on $\bp^{m ({m + \pdeg} \choose \pdeg - 1)}$
and the image of $C$ as the intersection of a Veronese variety
with a hyperplane. Then we can reduce inductively adding hypersurfaces
to the case of adding hyperplanes. However, this is accomplished
by the method used in Theorem ~\ref{thm:proj-one-point} in an
inductive manner as mentioned in Remark ~\ref{rem:proj-one-point-ind}.
\end{proof}

\ssec{Bounds for Arbitrary Divisors on Projective Space}
\begin{rem}
When $\langle f_1, \ldots, f_r \rangle = \langle x_0, \ldots, x_n \rangle$ (the irrelevant ideal), we may directly use the hypersurface method by considering $x_0, \ldots, x_n$ instead of the $f_i$'s. However, when $\sqrt{\langle f_1, \ldots, f_r \rangle} = \langle x_0, \ldots, x_n \rangle$ (or the saturation), we cannot obtain the same result from our modification of O'Dorney's method as seen in Example ~\ref{eg:radical}. Instead we must use ghost points and hypersurfaces.
\end{rem} \todo{Peter: What is the purpose of this remark?  The only method we have extends to ghost points so that this is satisfied.}

@@

This example further generalizes to the following situation: Suppose $D = \sum_{i=1}^{n} \frac{\num_i}{\den_i}D_i \in \di \bp^m \otimes \bq,$ where $\deg D_i = \bida_i$ and \todo{Peter: We need some condition that $D_i$'s be irreducible divisors or something.} $\deg D = \frac{1}{\lcm_{1 \leq i \leq n}(\den_i \cdot \bida_i)}$. Then, if $D_i = V(f_i)$ where all $f_i$ can be written as a polynomial function in $x_0, \ldots, x_{m-1},$ we have $R_D$ always has a generator in degree $\lcm(\den_i \cdot \bida_i)$.
\end{example}

@@

Consider $\frac{-1}{5}V(x_0^2 + x_1^2 + x_2^2) + \frac{1}{7}V(x_0^2 + x_1^2 + x_3^2) + \frac{1}{17}V(x_0^2 + x_2^2 + x_3^2) + \frac{1}{596}V(x_1^2 + x_2^2 + x_3^2)$. In degree $355216 = 5* 7 * 17 * 596$, this has dimension $10$. 
\todo{Clarify what this example means by the lattice approach}
However, we possibly use the lattice approach, as all lower degrees only have dimension 1. In fact, we may construct similar examples using any set of degree $n$ divisor with very small degree.
\end{example}
\todo{construct class of additional examples}
\todo{The example above needs to be changed because $x_0^2$ can be replaced by $2x_0$}

We now move on offer bounds on generators and relations of $R_D$ for a general divisor $D$ on $\bp^m$.

Let $D = \sum_{i=1}^n \alpha_i C_i$ for curves $C_i$ on $\mathbb{P}^m$ and let $f_i$ be homogeneous polynomials in $x_0, \ldots, x_m$ such that $C_i = v(f_i)$ for all $i$; suppose $\deg(f_i) = a_i$. 
\begin{lem} \label{P-m-generators}
$R_D$ is generated in degree at most $\sum_{i=1}^n \ell_i a_i$.
% and relations in degree at most  .\todo{insert numbers}
\end{lem}
\begin{proof}
Let 
\begin{align}\label{eqn:Sigma-def}
	\Sigma = \{(d, c_1, \ldots, c_n) \in \bz^{n+1} : c_i \geq - d \alpha_i, \; 1 \leq i \leq n, \text{ and } \sum_{i=1}^{n}k_i^j = 0, 1 \leq j \leq t\}.
\end{align}
We note that $\Sigma$ has extremal rays given by the lattice points 
\begin{equation}\label{defn:e_i-P-n}
	e_i = (\ell_i \bida_i, - \alpha_1 \ell_i \bida_i, \ldots -\alpha_{i-1} \ell_i \bida_i, \ell_i \sum_{j\ne i} \alpha_j \bida_j, -\alpha_{i+1} \ell_i \bida_i, \ldots, -\alpha_n, \ell_i \bida_i)
\end{equation}
for each $i\in \{1, \ldots n\}$.
Viewed inside of $R_D$, we have $e_i$ corresponding to
\begin{equation}\label{eqn:epilon-def-P-m}
	\epsilon_i = \prod_{j\ne i} (f_j)^{-\alpha_j \ell_i a_i} {f_i}^{\ell_i \sum_{j\ne i} \alpha_j a_j}\in R_D.
\end{equation}
Therefore, applying proposition \ref{prop:cone-generation}, we see $R_D$ is generated in degrees less than
\[
	\sum_{i=1}^n \ell_i a_i.
\]
\end{proof}

Let $x_1, \ldots x_r$ be the generators in degrees at most $\sum_{i=1}^n \ell_i \bida_i$ (given by lemma \ref{P-m-generators}), and let $\phi$ be the surjection $k[x_1, \ldots x_r] \to R_D$.  We can factor $\phi$ through 
\[
	\Sigma =  \langle u^d z_1^{c_1} \cdots z_n^{c_n} : c_i \in \mathbb{Z}, \; c_i \geq -d \alpha_i, \mbox{ and }\sum_{i=1}^{n} \bida_i c_i = \sum_{i=1}^{n} \bidb_i c_i. \rangle 
\]
by
\[
\begin{tikzcd}
\bk[x_1,\ldots, x_r] \ar {r}{\chi} & \bk[\Sigma] \ar {r}{\psi} & R_D \\
x_i \ar r & u^{d_i}z_1^{c_{i1}} \cdots z_n^{c_{in}} \ar r & u^{d_i}f_1^{c_{i1}} \cdots f_n^{c_{in}}.
\end{tikzcd}
\]

By Lemma \ref{lem:bound-ker-chi}, we can bound the degree of generation of $\ker(\chi)$ below
\[
	2 \sum_{i=1}^n \ell_i a_i.
\]
Finally, we calculate the degree of generation of $\psi$:

\begin{lem}\label{P-m-relations-psi}
$\ker(\psi)$ is generated in degree less than
\[
	\frac{\max_{1\le i \le n}(\bida_i)}{\deg(D)} +  \sum_{i=1}^n \ell_i a_i.
\]
\end{lem}
\begin{proof}
We first claim that $\psi$ there are $\beta_1, \ldots, \beta_n$ such that $\psi$ has relations of the form
\begin{equation}\label{eqn:relations-psi-P-n}
	(z_i - \beta_i)\prod_{j=1}^n {z_j}^{c_{j}}
\end{equation}

@@

To find such a $\beta$, we simply write $f_i$ as a 
polynomial in $f_1, \ldots f_3$. Furthermore, these
generate all relations, since they allow us to reduce any $\prod_{j =
1}^n f_j^{r_i}$ to a canonical form as given by Proposition \ref{prop:cone-generation}.

@@

\begin{thm}\label{P-m-generators/relations}
Let $D = \sum_{i=1}^n \alpha_i C_i$ for curves $C_i$ on $\mathbb{P}^m$.  Let $f_i$ be homogeneous polynomials in $x_0, \ldots, x_m$ such that $C_i = v(f_i)$ for all $i$; suppose $\deg(f_i) = a_i$. 
Then $R_D$ is generated in degree at most 
\[
	\sum_{i=1}^n \ell_i a_i
\]
with relations generated in degree at most
\[
	\max \left(2 \sum_{i=1}^n \ell_i a_i, \frac{\max_{1\le i \le n}(\bida_i)}{\deg(D)} + \sum_{i=1}^n \ell_i a_i \right).
\]
\end{thm}

\section{Canonical Rings on Hirzebruch surfaces}
Let $D=\sum_{i=1}^n \alpha_i C_i \in \mathbb{Q} \otimes \di(\hirz_
k)$ for some irreducible curves $C_1, \ldots C_n$ of $\hirz_k$, such
that $C_1, \ldots, C_4$ are independent curves with bi-degrees $(1,0)
, (1,0), (0,1)$, and $(0,1)$ respectively. Let $C_i = V(f_i)$ where
$f_i \in \mathscr{O}(\bida_i, \bidb_i)$.  Suppose $f_1$ and $f_2$ are independent linear
polynomials in $x_0$ and $x_1$ and $f_3$ and $f_4$ are independent
linear polynomials in $y_0$ and $y_1$.

@@

\begin{thm}\label{Hirzebruch-generators/relations}
Let $D = \sum_{i=1}^n \alpha_i C_i$ for curves $C_i$ on $\hirz_k$.  Let $C_i = V(f_i)$ where $f_i \in \mathscr{P}(a_i, b_i)$; further suppose $f_1$ and $f_2$ are independent linear
polynomials in $x_0$ and $x_1$ and $f_3$ and $f_4$ are independent
linear polynomials in $y_0$ and $y_1$.
Then $R_D$ is generated in degree less than 
\[
	\sigma = \sum_{i\in \Te(D)} \gcd(\bida_i, \bidb_i)\ell_i + \sum_{i\in \Tp(D) \atop
	j \in \Tm(D)} (\bida_i \bidb_j- \bida_j \bidb_i)\ell_{i,j}
\]
with relations generated in degree less than 
\[
	\max(2 \sigma, \sigma + \tau)
\]
where
\[
	\tau = \sigma
	+ \max( \max_{i\in S_=}(\ell_i \gcd(a_i, b_i)), \max_{i\in S_+ \atop j\in S_-}(\ell_{i,j} (\bida_i \bidb_j - \bida_j \bidb_i))).
\]
\end{thm}

@@

\section{Conclusion}
This paper extends the work of O'Dorney on $\mathbb{P}^1$ to $\mathbb{P}^m$ and Hirzebruch Surfaces, bounding the degrees of generation and relations of canonical rings of arbitrary divisors $D$.  The methods for deducing bounds on generators and relations of $\hirz_k$ applied similar ideas to O'Dorney's method on $\mathbb{P}^1$.

We can view $\hirz_0$ as $\mathbb{P}^1\times \mathbb{P}^1$ and more generally $\hirz_k$ as a $\mathbb{P}^1$ bundle over $\mathbb{P}^1$.  One might hope that similar ideas could be generalized to all bundles.  Specifically:
\begin{question}\label{qn:general-product/bundle}
Suppose $X$ and $Y$ are schemes.  Let $D$ be a divisor on $X\times Y$ or more generally on an $X$ bundles over $Y$.  Can we describe $D$ in terms of divisors $D_1, \ldots D_n$ on $X$ and $D_1', \ldots d_n'$ on $Y$.  Furthermore, can we find generators for $R_D$ and its ideal of relations using those of the $R_{D_i}$'s and $R_{D_j'}$'s?
\end{question}
An immediate obstruction to this is the fact that we critically used the fact that $\hirz_k$ has well-defined bi-degrees.  Unfortunately, in general $\Pic(X \times Y) \not \cong \Pic(X) \times \Pic(Y)$.

Alternatively, we can view $\hirz_1$ as $\mathbb{P}^2$ blown up at a point.  This suggests the following question:
\begin{question}\label{qn:general-blowup}
Let $X$ be a scheme.  Let $X'$ be the blowup of $X$ at a point $P$ with a divisor $D$.  First, is there a way to express $D$ in terms of divisors $D_1, \ldots, D_n$ on $X$?  Furthermore, is there a way to describe generators and relations of $R_D$ in terms of those of $R_{D_i}$ for all $i$.
\end{question}
Fortunately, the picard group of $X$ blown up at a point is the same as that of $X$ times an additional copy of $\bz$.  Perhaps analyzing the argument from \ref{Hirzebruch-generators/relations} in the case of $\hirz_k$ from the perspective of adding a copy of $\mathbb{Z}$ to the Picard group, as induced by the blowup, might suggest a general method for blowups.

>>>>

\ssec{Notation}
\begin{convention}
Notate a fractional (weil) divisor
\begin{align*}
	D = \sum_{i=1}^{n}\alpha_i D_i \in \di X \otimes \bq.
\end{align*}
where $\alpha_i \in \bq$ and $D_i \in \di X$ is an irreducible divisor.
\end{convention}

@@

\begin{proof}
By definition of $H^0(\bp^m,dD)$, functions satisfying conditions 
~\eqref{align:pm-span} lie in $H^0(\bp^m,dD)$. To complete the proof, it suffices to check functions satisfying conditions ~\eqref{align:pm-basis} form a basis of $H^0(\bp^m,dD)$. Note that there $\binom{m+ \lfloor dD \rfloor }{m}$ functions satisfying condition ~\eqref{align:pm-basis}. However we know $h^0(\bp^m,dD) = \binom{m+ \lfloor dD \rfloor }{m},$ so it suffices to show that those functions satisfying condition ~\eqref{align:pm-basis} are independent. This follows from the assumption that $f_0,\ldots f_n$ form a basis of degree 1 rational functions, and so degree $\lfloor d \deg D \rfloor $ monomials in $f_0,\ldots, f_n$ form a basis of degree $\lfloor d \deg D \rfloor $ rational functions.
\end{proof}

\ssec{Effective Divisors on Projective Space}
\label{ssec:proj-one-point}

In this subsection, we restrict attention to the case of effective divisors.

\begin{rem}
\label{rem:proj-veronese}
For any divisor $D = \sum_{i = 0}^{n} \alpha_i D_i$ on $\bp^m$
with each $D_i$ a hypersurface, we can use the Veronese mappings
to view $D$ as a divisor on $\bp^{m'}$ for some $m' \in \bz_{\geq 0}$
with $D = \sum_{i = 0}^{n} \alpha_i H_i$ where $H_i$ are hypersurfaces
in $\bp^{m'}$.
\end{rem}

@@

\begin{rem}
\label{rem:proj-one-point-ind}
Note that the method of proof for Theorem ~\ref{thm:proj-one-point}
can be immediately generalized to work inductively. If $D = D' + H$
for some hyperplane $H$, then we can use the proof above to show
the canonical ring $R_D$ is generated over $R_D'$ by the same
generators and with the same relations given in Theorem
~\ref{thm:proj-one-point}.
\end{rem}

\begin{thm}
\label{thm:proj-effective}
Let $X = \bp^m$ with $\{x_0, \ldots, x_m\}$ a coordinate system for
$\bp^m$. Let $D = \sum_{i = 1}^{n} \alpha_i D_i$, with $\alpha_i =
\frac{p_i}{q_i} \in \bq$ and each $D_i$ a hypersurface of
$\bp^m$, be an effective divisor on $\bp^m$.

@@

\begin{proof}
First, we apply Remark ~\ref{rem:proj-veronese} to reduce the
problem to letting each $H_i$ be a hyperplane of $\bp^m$.
Then, the proof follows from applying Remark
~\ref{rem:proj-one-point-ind} inductively on a one-hyperplane case
from Theorem ~\ref{thm:proj-one-point}.
\end{proof}

\ssec{Bounds for Arbitrary Divisors on Projective Space}

@@

This example further generalizes to the following situation: Suppose $D = \sum_{i=1}^{n} \frac{\num_i}{\den_i}D_i \in \di \bp^m \otimes \bq,$ where $\deg D_i = \bida_i$ and $\deg D = \frac{1}{\lcm_{1 \leq i \leq n}(\den_i \cdot \bida_i)}$. Then, if $D_i = V(f_i)$ where all $f_i$ can be written as a polynomial function in $x_0, \ldots, x_{m-1},$ we have $R_D$ always has a generator in degree $\lcm(\den_i \cdot \bida_i)$.
\end{example}

@@

Consider $\frac{-1}{5}V(x_0^2 + x_1^2 + x_2^2) + \frac{1}{7}V(x_0^2 + x_1^2 + x_3^2) + \frac{1}{17}V(x_0^2 + x_2^2 + x_3^2) - \frac{1}{596}V(x_1^2 + x_2^2 + x_3^2)$. In degree $355216 = 5 \cdot 7 \cdot 17 \cdot 596$, this has dimension $10$. 
\todo{Clarify what this example means by the lattice approach}
However, we possibly use the lattice approach, as all lower degrees only have dimension 1. In fact, we may construct similar examples using any set of degree $n$ divisor with very small degree.
\end{example}
\todo{construct class of additional examples}
\todo{The example above needs to be changed because $x_0^2$ can be replaced by $2x_0$}

We now offer bounds on generators and relations of $R_D$ for a general divisor $D$ on $\bp^m$.

Let $D = \sum_{i=1}^n \alpha_i D_i \in \bq \otimes \di \bp^m$ and let $f_i$ be homogeneous polynomials in $x_0, \ldots, x_m$ such that $D_i = V(f_i)$ for all $i$; suppose $\deg(f_i) = a_i$.
\todo{Aaron: Do we want to make the convention somewhere that $f_1,\ldots, f_{m+1}$ generate $O(1)$?}

\begin{lem} \label{P-m-generators}
$R_D$ is generated in degrees up to $\sum_{i=1}^n \ell_i a_i.$
% and relations in degree at most  .\todo{insert numbers}
\end{lem}
\begin{proof}
Let 
\begin{align}\label{eqn:Sigma-def}
	\Sigma = \left \{(d, c_1, \ldots, c_n) \in \bz^{n+1} : c_i \geq - d \alpha_i, \; 1 \leq i \leq n, \text{ and } \sum_{i=1}^{n} \ell_i \bida_i = 0 \right \}.
\end{align}
We note that $\Sigma$ has extremal rays given by the lattice points 
\begin{equation}\label{defn:e_i-P-n}
	e_i = \left(\ell_i \bida_i, - \alpha_1 \ell_i \bida_i, \ldots -\alpha_{i-1} \ell_i \bida_i, \ell_i \sum_{j\ne i} \alpha_j \bida_j, -\alpha_{i+1} \ell_i \bida_i, \ldots, -\alpha_n, \ell_i \bida_i \right)
\end{equation}
for each $i\in \{1, \ldots n\}$.
Therefore, applying proposition \ref{prop:cone-generation}, we see $R_D$ is generated in degrees less than
\[
	\sum_{i=1}^n \ell_i a_i.
\]
\end{proof}

\begin{rem}
\label{rem:pm-extremal-rays}
Note for future reference that the extremal rays of the cone $\Sigma$, as defined in the proof of Lemma ~\ref{P-m-generators} correspond to the elements
\begin{equation}\label{eqn:epilon-def-P-m}
	\epsilon_i = \prod_{j\ne i} (f_j)^{-\alpha_j \ell_i a_i} {f_i}^{\ell_i \sum_{j\ne i} \alpha_j a_j}\in R_D.
\end{equation}
\end{rem}

Let $x_1, \ldots x_r$ be the generators in degrees at most $\sum_{i=1}^n \ell_i \bida_i$ (given by lemma \ref{P-m-generators}), and let $\phi$ be the surjection $k[x_1, \ldots x_r] \to R_D$.  We can factor $\phi$ through the semigroup ring 
\[
	\bk[\Sigma] =  \langle u^d z_1^{c_1} \cdots z_n^{c_n} : c_i \in \mathbb{Z}, \; c_i \geq -d \alpha_i, \mbox{ and }\sum_{i=1}^{n} \bida_i c_i = \sum_{i=1}^{n} \bidb_i c_i \rangle. 
\]
by
\[
\begin{tikzcd}
\bk[x_1,\ldots, x_r] \ar {r}{\chi} & \bk[\Sigma] \ar {r}{\psi} & R_D \\
x_i \ar r & u^{d_i}z_1^{c_{i1}} \cdots z_n^{c_{in}} \ar r & u^{d_i}f_1^{c_{i1}} \cdots f_n^{c_{in}}.
\end{tikzcd}
\]

By Lemma \ref{lem:bound-ker-chi}, we can bound the degree of generation of $\ker(\chi)$ and we next calculate the degree of generation of $\ker \psi$:

\begin{lem}\label{P-m-relations-psi}
$\ker(\psi)$ is generated in degrees up to
\[
	\frac{\max_{1\le i \le n}(\bida_i)}{\deg(D)} +  \sum_{i=1}^n \ell_i a_i.
\]
\end{lem}
\begin{proof}
We first claim there exist $\beta_1, \ldots, \beta_n \in k[\Sigma]$ such that $\ker \psi$ is generated by
\begin{equation}\label{eqn:relations-psi-P-n}
	(z_i - \beta_i)\prod_{j=1}^n {z_j}^{c_{j}}
\end{equation}

@@

To find such a $\beta_i$, we simply write $f_i$ as a 
polynomial in $f_1, \ldots f_{m+1}$. Furthermore, these
generate all relations, since they allow us to reduce any $\prod_{j =
1}^n f_j^{r_i}$ to a canonical form as given by Proposition \ref{prop:cone-generation}.

@@

\begin{thm}\label{P-m-generators/relations}
Let $D = \sum_{i=1}^n \alpha_i D_i \in \bq \otimes \di \bp^m$. Let $f_i$ be homogeneous polynomials in $x_0, \ldots, x_m$ such that $D_i = V(f_i)$ for all $i$; suppose $\deg(f_i) = a_i$. 
Then $R_D$ is generated in degree at most 
\[
	\sum_{i=1}^n \ell_i a_i
\]
with relations generated in degree at most
\[
	\max \left(2 \sum_{i=1}^n \ell_i a_i, \frac{\max_{1\le i \le n}(\bida_i)}{\deg(D)} + \sum_{i=1}^n \ell_i a_i \right).
\]
\end{thm}

\section{Canonical Rings on Hirzebruch surfaces}
Let $D=\sum_{i=1}^n \alpha_i D_i \in \mathbb{Q} \otimes \di(\hirz_
k).$ such
that $D_1, \ldots, D_4$ are distinct divisors with bi-degrees $(1,0)
, (1,0), (0,1)$, and $(0,1)$ respectively. Let $D_i = V(f_i)$ where
$f_i \in \mathscr{O}(\bida_i, \bidb_i)$.  Suppose $f_1$ and $f_2$ are independent linear
polynomials in $x_0$ and $x_1$ and $f_3$ and $f_4$ are independent
linear polynomials in $y_0$ and $y_1$.

@@

\begin{thm}\label{Hirzebruch-generators/relations}
Let $D = \sum_{i=1}^n \alpha_i D_i$ for curves $D_i$ on $\hirz_k$.  Let $D_i = V(f_i)$ where $f_i \in \mathscr{P}(a_i, b_i)$; further suppose $f_1$ and $f_2$ are independent linear
polynomials in $x_0$ and $x_1$ and $f_3$ and $f_4$ are independent
linear polynomials in $y_0$ and $y_1$.
Then $R_D$ is generated in degree less than 
\[
	\sigma = \sum_{i\in \Te(D)} \gcd(\bida_i, \bidb_i)\ell_i + \sum_{i\in \Tp(D) \atop
	j \in \Tm(D)} (\bida_i \bidb_j- \bida_j \bidb_i)\ell_{i,j}
\]
with relations generated in degree less than 
\[
	\max(2 \sigma, \sigma + \tau)
\]
where
\[
	\tau = \sigma
	+ \max( \max_{i\in S_=}(\ell_i \gcd(a_i, b_i)), \max_{i\in S_+ \atop j\in S_-}(\ell_{i,j} (\bida_i \bidb_j - \bida_j \bidb_i))).
\]
\end{thm}

@@

\section{Conclusion}
This paper extends the work of O'Dorney on $\bp^1$ to $\bp^m$ and Hirzebruch Surfaces, bounding the degrees of generation and relations of canonical rings of arbitrary divisors $D$.  The methods for deducing bounds on generators and relations of $\hirz_k$ applied similar ideas to O'Dorney's method on $\bp^1$.

We can view $\hirz_0$ as $\bp^1\times \bp^1$ and more generally $\hirz_k$ as a $\bp^1$ bundle over $\bp^1$.  One might hope that similar ideas could be generalized to all bundles.  Specifically:
\begin{question}\label{qn:general-product/bundle}
Suppose $X$ and $Y$ are schemes.  Let $D$ be a divisor on $X\times Y$ or more generally on an $X$ bundles over $Y$.  Can we describe $D$ in terms of divisors $D_1, \ldots D_n$ on $X$ and $D_1', \ldots d_n'$ on $Y$.  Furthermore, can we find generators for $R_D$ and its ideal of relations using those of the $R_{D_i}$'s and $R_{D_j'}$'s?
\end{question}
An immediate obstruction to this is the fact that we critically used the fact that $\hirz_k$ has well-defined bi-degrees.  Unfortunately, in general $\Pic(X \times Y) \not \cong \Pic(X) \times \Pic(Y)$.

Alternatively, we can view $\hirz_1$ as $\bp^2$ blown up at a point.  This suggests the following question:
\begin{question}\label{qn:general-blowup}
Let $X$ be a scheme.  Let $X'$ be the blowup of $X$ at a point $P$ with a divisor $D$.  First, is there a way to express $D$ in terms of divisors $D_1, \ldots, D_n$ on $X$?  Furthermore, is there a way to describe generators and relations of $R_D$ in terms of those of $R_{D_i}$ for all $i$.
\end{question}
Fortunately, the picard group of $X$ blown up at a point is the same as that of $X$ times an additional copy of $\bz$.  Perhaps analyzing the argument from \ref{Hirzebruch-generators/relations} in the case of $\hirz_k$ from the perspective of adding a copy of $\mathbb{Z}$ to the Picard group, as induced by the blowup, might suggest a general method for blowups.
